---
title: <center>Visualizations of Set_E Subsets (Features)</center>
author: "Herman Bank"
output:
  html_document:
    df_print: paged
---

<br>

<h4>This program will do the following:

1. Separate the set_E subsets into individual .rds files and store in a local directory

2. Visualizations (violin plots) on RNA feature (nfeatures) and RNA count (nCount) for each subset.

3. Store the visualizations as png files in a local directory 

<br>

To enhance visual clarity and emphasize the significance of highly prevalent and most important genes, the violin plots data were filtered to a specified maximum count as a measure of quality control.

The RNA feature (nfeatures) plots ranged from 0 to 2500 with increments of 250, while the RNA count (nCount) plots ranged from 0 to 5000 with increments of 500.</h4>

<br>
Need to change the directory path, file path and output path to your local machine addresses
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# libraries
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(tidyverse)
library(ggplot2)

```


Creating a list of all set_E subsets
```{r}

# # Loading set_E
# set_E <- readRDS("~/TeamB/SeuratObjects/set_E.rds")
# 
# # Listing all the set_E subsets
# factor_levels <- levels(set_E@active.ident)
# print(factor_levels)


```


Separates the E subsets as individual files and stores in a local directory
```{r}

# #Name of all the subsets
# factors<- c("recur-M504", "recur-M519", "Wk2-M486", "Wk2-M488", "Wk10-M509", "Wk4-M482", "Wk4-M489", "Wk7-M502", "Wk0-M505", "Wk0-M514")
# 
# # Create a directory to store the subsets
# dir <- "~/TeamB/set_E_files"
# dir.create(dir)
# 
# # Loop through factors
# for (factor in factors) {
#   # Subset the data based on the factor
#   subset_data <- subset(set_E, orig.ident == factor)
#   subset_data[['percent.mt']] <- PercentageFeatureSet(subset_data, pattern = 'mt.')
# 
#   # Save the subset as an RDS file
#   file_name <- paste0(dir, "/", factor, ".rds")
#   saveRDS(subset_data, file = file_name)
# 
#  }
# 
# # Remove unnecessary data from memory
# remove('subset_data')
# remove('set_E')

```


Most peoples desktops and laptops do not have the memory to combine all 10 subsets as a list, therefore running 2 subsets at a time for the visualizations, then removing unneeded files from memory
```{r}

# #DONT HAVE ENOUGH MEMORY TO RUN
# # This is the code theat would be used if had suffiecient memory
# # Directory path
# directory_path <- "~/TeamB/set_E_files"
# 
# # Get the list of file names from the directory
# file_list <- list.files(directory_path, pattern = "\\.rds$", full.names = TRUE)
# 
# # Load the RDS files using lapply()
# setE_objects <- lapply(file_list, readRDS)


```


```{r}

output_dir <- "~/TeamB/set_E_Plots/"

recur_M504 <- readRDS("~/TeamB/set_E_files/recur-M504.rds")
recur_M519 <- readRDS("~/TeamB/set_E_files/recur-M519.rds")

set_list <- c(recur_M504, recur_M519)
set_names <- c("recur_M504", "recur_M519")
```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- set_names[i]
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
   # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nFeature_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    titleText <- "nCount_RNA"
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    
     titleText <- "percent.mt"
    plot <- VlnPlot(set, features = c("percent.mt"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 2)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_percent.mt.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
  })
}

# Removing unneeded files from memory
remove('set')
```


```{r}

# Removing unneeded files from memory
remove('set_list', 'recur_M504', 'recur_M519')

Wk2_M486 <- readRDS("~/TeamB/set_E_files/Wk2-M486.rds")
Wk2_M488 <- readRDS("~/TeamB/set_E_files/Wk2-M488.rds")

set_list <- c(Wk2_M486, Wk2_M488)
set_names <- c('Wk2_M486','Wk2_M488')

```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- set_names[i]
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250))  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
   # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nFeature_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    titleText <- "nCount_RNA"
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    
     titleText <- "percent.mt"
    plot <- VlnPlot(set, features = c("percent.mt"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 2)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
  })
}

# Removing unneeded files from memory
remove('set')
```


```{r}

# Removing unneeded files from memory
remove('set_list','Wk2_M486','Wk2_M488')

Wk4_M482 <- readRDS("~/TeamB/set_E_files/Wk4-M482.rds")
Wk4_M489 <- readRDS("~/TeamB/set_E_files/Wk4-M489.rds")

set_list <- c(Wk4_M482, Wk4_M489)
set_names <- c('Wk4_M482', 'Wk4_M489')

```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- set_names[i]
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250))  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
   # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nFeature_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    titleText <- "nCount_RNA"
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    
     titleText <- "percent.mt"
    plot <- VlnPlot(set, features = c("percent.mt"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 2)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
  })
}

# Removing unneeded files from memory
remove('set')
```


```{r}

# Removing unneeded files from memory
remove('set_list','Wk4_M482','Wk4_M489')

Wk7_M502 <- readRDS("~/TeamB/set_E_files/Wk7-M502.rds")
Wk10_M509 <- readRDS("~/TeamB/set_E_files/Wk10-M509.rds")

set_list <- c(Wk7_M502, Wk10_M509)
set_names <- c('Wk7_M502', 'Wk10_M509')


```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- set_names[i]
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250))  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
   # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nFeature_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    titleText <- "nCount_RNA"
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    
     titleText <- "percent.mt"
    plot <- VlnPlot(set, features = c("percent.mt"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 2)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
  })
}

# Removing unneeded files from memory
remove('set')
```


```{r}

# Removing unneeded files from memory
remove('set_list','Wk7_M502','Wk10_M509')

Wk0_M505 <- readRDS("~/TeamB/set_E_files/Wk0-M505.rds")
Wk0_M514 <- readRDS("~/TeamB/set_E_files/Wk0-M514.rds")

set_list <- c(Wk0_M505, Wk0_M514)
set_names <- c('Wk0_M505', 'Wk0_M514')

```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- set_names[i]
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250))  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
   # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nFeature_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    titleText <- "nCount_RNA"
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
    
    
     titleText <- "percent.mt"
    plot <- VlnPlot(set, features = c("percent.mt"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) +  scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 5)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    # Save the plot as an image file
    filename <- paste0(output_dir, set_name, "_nCount_RNA.png")
    ggsave(filename, plot, width = 6, height = 4, dpi = 300)
 
    # Print the plot
    print(plot)
  })
}

# Removing unneeded files from memory
remove('set')
```


```{r}

# Removing unneeded files from memory
remove('set_list','Wk0_M505', 'Wk0_M514')
```

