---
title: <center>RNA Feature and Count Visualiztions For All Sets</center>
author: "Herman Bank"
output:
  html_document:
    df_print: paged
---
<br>

<h4>This program will do the following:

1. Remove the control groups from set A

2. For all sets, calculate mitochondrial QC metrics with the PercentageFeatureSet() function, which calculates the percentage of counts originating from a set of features

3. Visualizations (violin plots) on RNA feature (nfeatures) and RNA count (nCount) for each set 

<br>

To enhance visual clarity and emphasize the significance of highly prevalent and most important genes, the violin plots data were filtered to a specified maximum count as a measure of quality control.

The RNA feature (nfeatures) plots ranged from 0 to 2500 with increments of 250, while the RNA count (nCount) plots ranged from 0 to 5000 with increments of 500.</h4>

<br>
Need to change the directory path, file path and output path to your local machine addresses
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# libraries
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(tidyverse)
library(ggplot2)
```


```{r}
set_A <- readRDS("~/TeamB/SeuratObjects/set_A.rds")
set_B <- readRDS("~/TeamB/SeuratObjects/set_B.rds")
set_C <- readRDS("~/TeamB/SeuratObjects/set_C.rds")
set_D <- readRDS("~/TeamB/SeuratObjects/set_D.rds")
set_E <- readRDS("~/TeamB/SeuratObjects/set_E.rds")
```


```{r}

 # Removing the control groups from set_A
set_A <- subset(set_A, orig.ident == c("SHAM-M505", "SHAM-M514"))
levels(set_A$orig.ident)
set_A$orig.ident <- droplevels(set_A$orig.ident)

```

```{r}

set_A[["percent.mt"]] <- PercentageFeatureSet(set_A, pattern = "^MT-")
set_B[["percent.mt"]] <- PercentageFeatureSet(set_B, pattern = "^MT-")
set_C[["percent.mt"]] <- PercentageFeatureSet(set_C, pattern = "^MT-")
set_D[["percent.mt"]] <- PercentageFeatureSet(set_D, pattern = "^MT-")
set_E[["percent.mt"]] <- PercentageFeatureSet(set_E, pattern = "^MT-")

```


Since the mitochondria percentage is the same for all factors in all sets, will ignore percent.mt
```{r}

set_list <- c(set_A,set_B,set_C,set_D,set_E)

# for (set in set_list){
#   suppressWarnings({
#     # Visualize QC metrics as a violin plot
#     plot <- VlnPlot(set, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
#     # Print the plot
#     print(plot)
#   })
# }
# 
# # Removing nuneccary data from memory
# remove('set')

```


```{r}

for (i in seq_along(set_list)) {
  set <- set_list[[i]]
  set_name <- paste0("Set ", LETTERS[i])
  titleText <- "nFeature RNA"
  suppressWarnings({
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nFeature_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) + ggtitle(paste(set_name, titleText)) + scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 2500), breaks = seq(0, 2500, by = 250)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
   
    # Print the plot
    print(plot)
    
    
  titleText <- "nCount RNA"
    # Visualize QC metrics as a violin plot
    plot <- VlnPlot(set, features = c("nCount_RNA"), pt.size = 0) +  theme(legend.position = "none") + ggtitle(paste(set_name, titleText)) + ggtitle(paste(set_name, titleText)) + scale_y_continuous(labels = scales::comma)
    plot <- plot + scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500)) + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
 
    # Print the plot
    print(plot)
  })
}


```



```{r}
remove('set','set_list', 'set_A', 'set_B','set_C','set_D','set_E')
```









